\begin{abstract}
Parameter-efficient fine-tuning (PEFT) methods such as LoRA, adapters, and BitFit have become standard tools for adapting large pre-trained models in natural language processing and computer vision. However, it remains unclear whether the same PEFT rankings transfer to geospatial foundation models, which differ in both architecture and input structure, and whether benchmark rankings established on curated public datasets reproduce in operational GIS deployments. In this work, we present a systematic benchmark of PEFT methods on Prithvi-100M under heterogeneous input settings, paired with a production validation on a real-world land-use/land-cover (LULC) workflow. We evaluate seven adaptation strategies across five modality configurations on EuroSAT, validate the main findings on BigEarthNet-S2, extend the benchmark to semantic segmentation on LandCover.ai, and re-instantiate the protocol on a 35{,}920-patch Chinese-domestic high-resolution corpus collected over Linhe County, Inner Mongolia, using Esri 2022 global LULC as the supervisory signal. In total, the study covers 109 public-benchmark experiments plus an additional production-deployment set, including a full fine-tuning baseline, a diagnostic LoRA split-QKV ablation, a dense-prediction replication, and a label-quality control study.

Our results reveal four consistent findings. First, LoRA provides virtually no gain over linear probing across both EuroSAT and BigEarthNet-S2. A diagnostic ablation shows that this failure is caused by two compounding factors: a fused-QKV implementation trap in standard PyTorch attention layers, and the fact that low-rank adaptation remains suboptimal even after Q/K/V are correctly exposed. Second, Houlsby adapters consistently dominate all other PEFT methods, reaching 94.5\% of full fine-tuning performance on EuroSAT while using only 1.4\% of the trainable parameters. Third, modality selection matters more than PEFT choice: multi-spectral inputs consistently outperform RGB subsets across methods and datasets. The LandCover.ai extension confirms that the Houlsby $\gg$ LoRA $\approx$ Linear~Probe ranking is not specific to classification: on dense prediction, the Houlsby advantage widens to $+155\%$ relative mIoU, while LoRA still collapses to the linear-probe baseline to the fourth decimal place. Fourth, the production deployment on Linhe reproduces the public-benchmark ranking under a geographic scene-level split protocol (Houlsby $+0.116$ mIoU over linear probing, $+65.5\%$ relative; LoRA collapses to within 0.0004 mIoU of linear probing, reproducing the fused-QKV failure mode across all four datasets), and a paired label-quality control on the same imagery shows that the PEFT delta is mediated by label fidelity: synthetic brightness-thresholded weak labels collapse the delta to $+0.017$, while real Esri labels recover the $+0.116$ gap.

These results suggest that PEFT intuitions from mainstream NLP and vision cannot be transferred directly to geospatial foundation models without validation, and that single-dataset PEFT comparisons can mislead practitioners about expected operating points. Our benchmark provides both empirical guidance for practitioners and a reproducible experimental framework, released together with the AlphaEarth-System platform that orchestrates the Linhe pipeline.
\end{abstract}
