\section{Conclusion}

We presented a systematic evaluation of PEFT methods for adapting Prithvi-100M to heterogeneous geospatial inputs. Across 100 total experiments spanning EuroSAT, BigEarthNet-S2, a full fine-tuning ceiling, and a diagnostic LoRA ablation, three conclusions emerge clearly. First, PEFT rankings from NLP and conventional vision do not transfer directly to geospatial foundation models. Second, Houlsby adapters provide the strongest and most reliable adaptation performance across datasets, tasks, and modality configurations. Third, modality selection has a larger practical impact than PEFT choice, especially when moving from reduced-band inputs to richer multi-spectral configurations.

The study also contributes a deeper diagnosis of LoRA failure in GeoFMs. Rather than a single negative result, we show a two-stage explanation: standard implementations miss fused QKV projections, and corrected low-rank adaptation remains inferior to adapter-based methods. This finding highlights the importance of architecture-aware PEFT evaluation in remote sensing.

Our work has several limitations. We focus on a single backbone, Prithvi-100M, and two benchmark datasets. Although the cross-dataset consistency is strong, further validation on additional geospatial foundation models and task types would strengthen the generality of the conclusions. Future work should also examine whether the same PEFT ranking holds for segmentation, regression, and temporal forecasting tasks.

Despite these limitations, we believe this benchmark provides a useful empirical foundation for the next stage of GeoFM research. By clarifying which PEFT methods work, which fail, and why, it offers both a practical guide for deployment and a methodological reference for future adaptation studies.
