\section{Discussion}
\label{sec:discussion}

\subsection{Why Does Low-Rank Adaptation Underperform on Prithvi?}
Our split-QKV ablation isolates two effects. The \emph{implementation trap} reflects a property of PyTorch's \texttt{nn.MultiheadAttention}, in which query, key, and value weights are packed into a single \texttt{in\_proj\_weight} tensor. A naive application of LoRA to named \texttt{nn.Linear} children therefore touches only \texttt{out\_proj}, leaving the attention projections effectively frozen. This is a silent failure mode: training proceeds, loss decreases, and accuracy settles only slightly above linear probing, which can easily be misread as a genuine negative result.

Once Q, K, and V are explicitly exposed, split-QKV LoRA improves EuroSAT s2\_full from 0.658 to 0.707 but remains below both BitFit (0.702) and Houlsby (0.821). We attribute this residual gap to two factors. First, the masked autoencoding objective~\cite{he2022mae} used by Prithvi produces features whose task-relevant directions are not well captured by low-rank updates to attention alone; richer adaptation (bottleneck MLPs) appears to be required. Second, the frozen backbone restricts LoRA to reshaping existing attention patterns, whereas Houlsby adapters introduce new non-linear capacity inside each block. This is consistent with the unified view of PEFT proposed in prior work~\cite{he2022unified}, which predicts that methods with both extra capacity and non-linearity can dominate under sufficiently different pre-training distributions. A follow-up rank-sensitivity ablation strengthens this interpretation: increasing split-QKV LoRA from $r=4$ to $r=32$ leaves EuroSAT s2\_full accuracy essentially flat at 0.706--0.709, ruling out the simple explanation that the negative result is caused by an underpowered rank choice alone.

\subsection{When Does Input-Stage Adaptation Help?}
GeoAdapter v2 places a small residual module before the patch embedding. Conceptually, this is attractive for modality bridging, but the empirical picture is regime-dependent. On EuroSAT, where the input is 13-band Sentinel-2 close to Prithvi's HLS pretraining modality, GeoAdapter underperforms linear probing on every modality configuration: the input-stage residual has limited new information to add, and the small parameter budget is insufficient to jointly learn modality alignment and useful feature transformation. On the Linhe production corpus (Section~\ref{sec:linhe}), where the input is 3-band RGB padded with zeros into the HLS template --- a substantial modality shift --- GeoAdapter recovers $+0.087$ mIoU over linear probing at only 4{,}756 parameters, second only to Houlsby's bottleneck adapters. We therefore refine the negative result: input-stage adaptation is fragile when the modality shift is small and the patch embedding is already well-tuned, but becomes useful as the input drifts further from the backbone's pretraining distribution. The recommendation is to treat GeoAdapter as a deployment-time decision: profile the modality gap first, then decide. Two underlying mechanisms still bound its potential. First, the module lies on the critical gradient path to a frozen backbone, so any early-stage misadaptation propagates unchallenged through 12 transformer blocks. Second, the adapter has no clear pre-training prior, unlike bottleneck adapters that operate on already well-structured hidden states. Future modality bridging work should combine input-stage and backbone-level adaptation rather than rely on either alone.

\subsection{Architecture-Awareness as a Methodological Principle}
The fused-QKV trap has an uncomfortable implication: published PEFT comparisons on Prithvi-like backbones may silently conflate \emph{method failure} with \emph{implementation failure}. We believe this motivates a broader methodological recommendation. Before reporting negative results for PEFT methods on any new architecture, authors should report (i) exactly which parameter tensors were registered as trainable, (ii) whether the attention projections are fused, and (iii) whether a diagnostic split variant was attempted. Our study adopts this protocol and we encourage its adoption for future GeoFM benchmarks.

\subsection{Practical Guidance and Limitations}
For practitioners adapting Prithvi-100M to heterogeneous inputs, the safest default is Houlsby adapters with bottleneck dimension 64, which reach 94.5\% of the full fine-tuning ceiling on EuroSAT s2\_full while training 1.4\% of parameters. When spectral richness is available, preserving it yields larger returns than switching between PEFT methods: a linear probe on s2\_full outperforms sophisticated adaptation on sar\_only. LoRA should not be used on fused-QKV backbones without explicit Q/K/V exposure and, even then, should be justified empirically against adapter baselines.

Our study has three limitations. First, we evaluate a single backbone, Prithvi-100M; extension to SatMAE~\cite{cong2022satmae}, Scale-MAE~\cite{reed2023scalemae}, and SpectralGPT~\cite{hong2024spectralgpt} would strengthen the generality of conclusions. Second, although we now cover classification (single-label on EuroSAT, multi-label on BigEarthNet-S2) and dense prediction (semantic segmentation on LandCover.ai), regression and temporal forecasting remain open. Third, our BigEarthNet-S2 subsample (10K/5K) is chosen for Colab-scale tractability and may under-represent the long tail of rare classes; full-scale replication is an obvious next step.
