\section{Introduction}

Geospatial foundation models (GeoFMs) have recently emerged as a promising paradigm for Earth observation analysis. Models such as Prithvi-100M~\cite{jakubik2023prithvi}, SatMAE~\cite{cong2022satmae}, Scale-MAE~\cite{reed2023scalemae}, and related self-supervised transformers~\cite{hong2024spectralgpt} suggest that large-scale pre-training on satellite imagery can yield transferable representations for downstream remote sensing tasks. At the same time, practitioners increasingly face adaptation problems that differ substantially from the homogeneous settings used during pre-training. Real-world inputs may contain different band subsets, missing channels, alternative sensor combinations, or strongly out-of-distribution modality configurations.

In mainstream natural language processing and computer vision, parameter-efficient fine-tuning (PEFT) methods are widely used to adapt large pre-trained models at low cost. LoRA~\cite{hu2022lora}, adapters~\cite{houlsby2019adapter}, BitFit~\cite{zaken2022bitfit}, and prompt-style tuning methods~\cite{lester2021prompt,jia2022vpt} have become standard choices because they often approach full fine-tuning performance with far fewer trainable parameters. This success naturally raises the question of whether the same PEFT rankings hold for geospatial foundation models.

We argue that this assumption is non-trivial for at least three reasons. First, geospatial foundation models often operate on heterogeneous multi-band inputs rather than fixed RGB inputs. Second, remote sensing pre-training objectives, especially masked autoencoding on spectral-spatial data, differ from the language modeling and image classification setups where PEFT methods are most established. Third, implementation details that are harmless in conventional vision backbones can become critical in GeoFMs, especially when architectural components such as fused attention projections interact with adaptation layers.

Despite the practical importance of this problem, current literature lacks a systematic benchmark of PEFT methods under heterogeneous geospatial input settings. Existing studies often evaluate only one or two adaptation strategies, focus on a single downstream task, or treat PEFT as an implementation detail rather than the object of study. As a result, practitioners still lack clear guidance on which PEFT method to choose when adapting a model such as Prithvi-100M to non-standard remote sensing inputs.

A second, related gap is that PEFT comparisons are almost always reported on curated public datasets. It is rarely demonstrated that the same ranking reproduces on operational GIS pipelines, where imagery is heterogeneous in sensor and resolution, labels are derived from third-party products of varying quality, and validation splits must respect spatial structure to prevent leakage. The absence of such a demonstration is a real obstacle for practitioners deciding whether to ship a PEFT-adapted GeoFM into a production workflow.

This paper addresses both gaps through a controlled empirical study on Prithvi-100M, paired with a production validation. We benchmark seven adaptation strategies across five modality configurations on EuroSAT, validate the main findings on BigEarthNet-S2, extend the protocol to semantic segmentation on LandCover.ai, and re-instantiate the resulting five-method comparison on a 35{,}920-patch Linhe County corpus collected from 203 Chinese-domestic high-resolution scenes (0.42--4.29\,m GSD, 22 satellite families) with Esri 2022 global LULC as labels. Beyond reporting final scores, we include a full fine-tuning ceiling, a diagnostic split-QKV LoRA ablation, and a paired label-quality control study on the production corpus. Across 109 public-benchmark experiments plus the Linhe deployment, we show that PEFT rankings from mainstream NLP and vision do not transfer directly to geospatial foundation models, and that benchmark rankings reproduce on production geography under a geographic-split protocol.

Our main contributions are as follows:
\begin{enumerate}[leftmargin=1.5em]
    \item We present the first systematic benchmark of PEFT methods on Prithvi-100M under heterogeneous input settings, covering 109 public-benchmark experiments across EuroSAT, BigEarthNet-S2, and LandCover.ai.
    \item We show that LoRA fails on Prithvi-100M for two compounding reasons: naive implementations miss fused QKV projections, and corrected low-rank adaptation remains suboptimal even after this issue is fixed.
    \item We demonstrate that Houlsby adapters consistently dominate all other PEFT methods across all four datasets, achieving 94.5\% of full fine-tuning performance with only 1.4\% of the trainable parameters on EuroSAT and a $+155\%$ relative mIoU gain on dense prediction.
    \item We confirm across datasets that modality selection has a larger practical impact than PEFT selection: richer spectral inputs consistently outperform reduced-band alternatives.
    \item We provide a negative result for standalone input-stage adaptation (GeoAdapter), clarifying its failure modes and limitations.
    \item We validate the benchmark ranking on an operational Linhe County LULC deployment under a geographic scene-level split, recovering Houlsby $+0.116$ mIoU over linear probing on production imagery, and provide a paired label-quality control study showing that the PEFT delta tracks label fidelity.
    \item We release a reproducible Geo-MLOps benchmarking framework together with the AlphaEarth-System platform that orchestrates the Linhe deployment, including its data pipeline, training engine, and web dashboard.
\end{enumerate}
